---
title: "ClimateVariabilityAndCatchability_filtered"
format: docx
---
### Revised Approach
This basically replicates the work done in Matlab, but restricts the analyses to 
grid cells that have at least three vessels' activity over the period of interest.
This threshold does three things:  
1. Focuses the analysis on waters that are actually being fished, thereby avoiding
strong signals at the periphery of the geographic area of interest.  
2. Sharpens this focus to waters that were fished more than once and by more than
a single operator.  
3. Renders the results non-confidential.

Worth noting, this sums monthly vessels across all months.  So, it's theoretically
possible that there could be a grid cell that's visited once a month for three months
by the same vessel.  To avoid this, we could make a separate `TotalVessels.nc` 
file that sums across all months instead of across each month.

```{r, echo = FALSE, message = FALSE}
# Set up workspace
library(tidyverse, quietly = TRUE)
library(tidync)
library(here)
```

```{r, echo = FALSE}
### Load data
# Thanks to phind.com for help with the netCDF stuff, and to Johanna for drawing
# my attention to the tidync package to begin with

# Load climate indices
ONI <- read_csv(here("ClimateIndices", "ONI_withPhases.csv"), show_col_types = FALSE)
PDO <- read_csv(here("ClimateIndices", "PDO.csv"), show_col_types = FALSE)
NPGO <- read_csv(here("ClimateIndices", "NPGO.csv"), show_col_types = FALSE)

# Create a tidync objects
Catchability_nc <- tidync(here("OceanData", "BigeyeCatchability.nc"))
GODAS_nc <- tidync(here("OceanData", "GODAS_1deg.nc"))
O2_nc <- tidync(here("OceanData", "O2_2mlpl_depth_1deg_noInterp.nc"))
Vessels_nc <- tidync(here("FisheryData", "TotalVessels.nc"))

# Extract variables
# Retaining netCDF dimensions
Catchability_a <- Catchability_nc |>
  hyper_array()
GODAS_a <- GODAS_nc |>
  hyper_array()
O2_a <- O2_nc |>
  hyper_array()
Vessels_a <- Vessels_nc |>
  hyper_array()
# In a tidy-compatible format (not used)
# Catchability_t <- Catchability_nc |>
#   hyper_tibble()
# GODAS_t <- GODAS_nc |>
#   hyper_tibble()
# O2_t <- O2_nc |>
#   hyper_tibble()
# Vessels_t <- Vessels_nc |>
#   hyper_tibble()

GODAS_Depth <- dimnames(GODAS_a$GODAS_REGRID)[3]
Depth <- as.numeric(GODAS_Depth$LEV1_31)
```

```{r, echo = FALSE}
### Wrangle data
# Limit climate indices to our period of interest: 1995–2024
ONI_9524 <- ONI |>
  filter(YR >= 1995 & YR <= 2024) |>
  select(ONI)
PDO_9524 <- PDO |>
  filter(Year >= 1995 & Year <= 2024) |>
  select(PDO)
NPGO_9524 <- NPGO |>
  filter(YEAR >= 1995 & YEAR <= 2024) |>
  select(NPGO)

# Clean up
rm(ONI, PDO, NPGO)

# Filter so that only cells with >=3 vessels are used
TotalVessels <- apply(Vessels_a$`Total Vessels`, c(1,2), sum, na.rm = TRUE)
for (r in seq(1, dim(TotalVessels)[1], 1)) {
  for (c in seq(1, dim(TotalVessels)[2], 1)) {
    if (TotalVessels[r,c] < 3) {
      Catchability_a$Catchability[r,c,] <- NA
      O2_a$GLORYS_O2_2MLPL_REGRID[r,c,] <- NA
      GODAS_a$GODAS_REGRID[r,c,,] <- NA
    }
  }
}
rm (r, c)
```

```{r, echo = FALSE, warning = FALSE}
# NOTE: The test for Spearman rank correlations result in a LOT of warnings about 
# ties, or equal values that can't be ranked.  Suppressing them.

### Correlations
# Empty arrays to fill
Catchability_Spearman <- array(NA, 
                               dim = c(dim(Catchability_a$Catchability)[1], 
                                       dim(Catchability_a$Catchability)[2], 
                                       3))
Catchability_Pearson <- array(NA, 
                               dim = c(dim(Catchability_a$Catchability)[1], 
                                       dim(Catchability_a$Catchability)[2], 
                                       3))

O2_Spearman <- array(NA, 
                     dim = c(dim(O2_a$GLORYS_O2_2MLPL_REGRID)[1],
                             dim(O2_a$GLORYS_O2_2MLPL_REGRID)[2],
                             3))
O2_Pearson <- array(NA, 
                    dim = c(dim(O2_a$GLORYS_O2_2MLPL_REGRID)[1], 
                            dim(O2_a$GLORYS_O2_2MLPL_REGRID)[2], 
                            3))

D8_Spearman <- array(NA, 
                     dim = c(dim(GODAS_a$GODAS_REGRID)[1],
                             dim(GODAS_a$GODAS_REGRID)[2],
                             3))
D8_Pearson <- array(NA, 
                    dim = c(dim(GODAS_a$GODAS_REGRID)[1], 
                            dim(GODAS_a$GODAS_REGRID)[2], 
                            3))

T814_Spearman <- array(NA, 
                     dim = c(dim(GODAS_a$GODAS_REGRID)[1],
                             dim(GODAS_a$GODAS_REGRID)[2],
                             3))
T814_Pearson <- array(NA, 
                    dim = c(dim(GODAS_a$GODAS_REGRID)[1], 
                            dim(GODAS_a$GODAS_REGRID)[2], 
                            3))

for (r in seq(1, dim(Catchability_a$Catchability)[1])) {
  for (c in seq(1, dim(Catchability_a$Catchability)[2])) {
    # Filtering for 3+ vessels introduces cells that are all NaNs.
    ### Catchability
    if (any(!is.na(Catchability_a$Catchability[r,c,]))) {
      # ONI
      Catch_Spearman_ONI <- cor.test(Catchability_a$Catchability[r,c,], ONI_9524$ONI,
                                     use = "pairwise.complete.obs", 
                                     method = "spearman")
      
      Catch_Pearson_ONI <- cor.test(Catchability_a$Catchability[r,c,], ONI_9524$ONI,
                                    use = "pairwise.complete.obs", 
                                    method = "pearson")
      
      if (Catch_Spearman_ONI$p.value < 0.05) {
        Catchability_Spearman[r,c,1] = Catch_Spearman_ONI$estimate
      }
      
      if (Catch_Pearson_ONI$p.value < 0.05) {
        Catchability_Pearson[r,c,1] = Catch_Pearson_ONI$estimate
      }
      
      # PDO
      Catch_Spearman_PDO <- cor.test(Catchability_a$Catchability[r,c,], PDO_9524$PDO,
                                     use = "pairwise.complete.obs", 
                                     method = "spearman")
      
      Catch_Pearson_PDO <- cor.test(Catchability_a$Catchability[r,c,], PDO_9524$PDO,
                                    use = "pairwise.complete.obs", 
                                    method = "pearson")
      
      if (Catch_Spearman_PDO$p.value < 0.05) {
        Catchability_Spearman[r,c,2] = Catch_Spearman_PDO$estimate
      }
      
      if (Catch_Pearson_PDO$p.value < 0.05) {
        Catchability_Pearson[r,c,2] = Catch_Pearson_PDO$estimate
      }
      
      # NPGO
      Catch_Spearman_NPGO <- cor.test(Catchability_a$Catchability[r,c,], NPGO_9524$NPGO,
                                      use = "pairwise.complete.obs", 
                                      method = "spearman")
      
      Catch_Pearson_NPGO <- cor.test(Catchability_a$Catchability[r,c,], NPGO_9524$NPGO,
                                     use = "pairwise.complete.obs", 
                                     method = "pearson")
      
      if (Catch_Spearman_NPGO$p.value < 0.05) {
        Catchability_Spearman[r,c,3] = Catch_Spearman_NPGO$estimate
      }
      
      if (Catch_Pearson_NPGO$p.value < 0.05) {
        Catchability_Pearson[r,c,3] = Catch_Pearson_NPGO$estimate
      }
    }
    
    ### Depth of 2 mlpl O2
    if (any(!is.na(O2_a$GLORYS_O2_2MLPL_REGRID[r,c,]))) {
      # ONI
      O2_Spearman_ONI <- cor.test(O2_a$GLORYS_O2_2MLPL_REGRID[r,c,], ONI_9524$ONI,
                                  use = "pairwise.complete.obs", 
                                  method = "spearman")
      
      O2_Pearson_ONI <- cor.test(O2_a$GLORYS_O2_2MLPL_REGRID[r,c,], ONI_9524$ONI,
                                 use = "pairwise.complete.obs", 
                                 method = "pearson")
      
      if (O2_Spearman_ONI$p.value < 0.05) {
        O2_Spearman[r,c,1] = O2_Spearman_ONI$estimate
      }
      
      if (O2_Pearson_ONI$p.value < 0.05) {
        O2_Pearson[r,c,1] = O2_Pearson_ONI$estimate
      }
      
      # PDO
      O2_Spearman_PDO <- cor.test(O2_a$GLORYS_O2_2MLPL_REGRID[r,c,], PDO_9524$PDO,
                                  use = "pairwise.complete.obs", 
                                  method = "spearman")
      
      O2_Pearson_PDO <- cor.test(O2_a$GLORYS_O2_2MLPL_REGRID[r,c,], PDO_9524$PDO,
                                 use = "pairwise.complete.obs", 
                                 method = "pearson")
      
      if (O2_Spearman_PDO$p.value < 0.05) {
        O2_Spearman[r,c,2] = O2_Spearman_PDO$estimate
      }
      
      if (O2_Pearson_PDO$p.value < 0.05) {
        O2_Pearson[r,c,2] = O2_Pearson_PDO$estimate
      }
      
      # NPGO
      O2_Spearman_NPGO <- cor.test(O2_a$GLORYS_O2_2MLPL_REGRID[r,c,], NPGO_9524$NPGO,
                                   use = "pairwise.complete.obs", 
                                   method = "spearman")
      
      O2_Pearson_NPGO <- cor.test(O2_a$GLORYS_O2_2MLPL_REGRID[r,c,], NPGO_9524$NPGO,
                                  use = "pairwise.complete.obs", 
                                  method = "pearson")
      
      if (O2_Spearman_NPGO$p.value < 0.05) {
        O2_Spearman[r,c,3] = O2_Spearman_NPGO$estimate
      }
      
      if (O2_Pearson_NPGO$p.value < 0.05) {
        O2_Pearson[r,c,3] = O2_Pearson_NPGO$estimate
      }
    }
    
    ### Depth of 8 deg C water and thickness of 8-14 deg C water
    if(any(!is.na(GODAS_a$GODAS_REGRID[r,c,,]))) {
      # Empty matrices to fill
      Depth8 <- matrix(NA, nrow = 360)
      Thick814 <- matrix(NA, nrow = 360)
      for (m in seq(1, 360, 1)) {
        Z8_idx <- which(GODAS_a$GODAS_REGRID[r,c,,m] >= 8)
        Z14_idx <- which(GODAS_a$GODAS_REGRID[r,c,,m] <= 14)
        
        Depth8[m] <- Depth[max(Z8_idx)]
        Thick814[m] <- Depth[max(Z8_idx)] - Depth[min(Z14_idx)]
      }
      rm(m, Z8_idx, Z14_idx)
      
      # ONI
      D8_Spearman_ONI <- cor.test(Depth8[,1], ONI_9524$ONI,
                                  use = "pairwise.complete.obs", 
                                  method = "spearman")
      
      D8_Pearson_ONI <- cor.test(Depth8, ONI_9524$ONI,
                                 use = "pairwise.complete.obs", 
                                 method = "pearson")
      
      if (!is.na(D8_Spearman_ONI$p.value) & D8_Spearman_ONI$p.value < 0.05) {
        D8_Spearman[r,c,1] = D8_Spearman_ONI$estimate
      }
      
      if (!is.na(D8_Pearson_ONI$p.value) & D8_Pearson_ONI$p.value < 0.05) {
        D8_Pearson[r,c,1] = D8_Pearson_ONI$estimate
      }
      
      # PDO
      D8_Spearman_PDO <- cor.test(Depth8, PDO_9524$PDO,
                                  use = "pairwise.complete.obs", 
                                  method = "spearman")
      
      D8_Pearson_PDO <- cor.test(Depth8, PDO_9524$PDO,
                                 use = "pairwise.complete.obs", 
                                 method = "pearson")
      
      if (!is.na(D8_Spearman_PDO$p.value) & D8_Spearman_PDO$p.value < 0.05) {
        D8_Spearman[r,c,2] = D8_Spearman_PDO$estimate
      }
      
      if (!is.na(D8_Pearson_PDO$p.value) & D8_Pearson_PDO$p.value < 0.05) {
        D8_Pearson[r,c,2] = D8_Pearson_PDO$estimate
      }
      
      # NPGO
      D8_Spearman_NPGO <- cor.test(Depth8, NPGO_9524$NPGO,
                                   use = "pairwise.complete.obs", 
                                   method = "spearman")
      
      D8_Pearson_NPGO <- cor.test(Depth8, NPGO_9524$NPGO,
                                  use = "pairwise.complete.obs", 
                                  method = "pearson")
      
      if (!is.na(D8_Spearman_NPGO$p.value) & D8_Spearman_NPGO$p.value < 0.05) {
        D8_Spearman[r,c,3] = D8_Spearman_NPGO$estimate
      }
      
      if (!is.na(D8_Pearson_NPGO$p.value) & D8_Pearson_NPGO$p.value < 0.05) {
        D8_Pearson[r,c,3] = D8_Pearson_NPGO$estimate
      }
      
      # ONI
      T814_Spearman_ONI <- cor.test(Thick814, ONI_9524$ONI,
                                    use = "pairwise.complete.obs", 
                                    method = "spearman")
      
      T814_Pearson_ONI <- cor.test(Thick814, ONI_9524$ONI,
                                   use = "pairwise.complete.obs", 
                                   method = "pearson")
      
      if (T814_Spearman_ONI$p.value < 0.05) {
        T814_Spearman[r,c,1] = T814_Spearman_ONI$estimate
      }
      
      if (T814_Pearson_ONI$p.value < 0.05) {
        T814_Pearson[r,c,1] = T814_Pearson_ONI$estimate
      }
      
      # PDO
      T814_Spearman_PDO <- cor.test(Thick814, PDO_9524$PDO,
                                    use = "pairwise.complete.obs", 
                                    method = "spearman")
      
      T814_Pearson_PDO <- cor.test(Thick814, PDO_9524$PDO,
                                   use = "pairwise.complete.obs", 
                                   method = "pearson")
      
      if (T814_Spearman_PDO$p.value < 0.05) {
        T814_Spearman[r,c,2] = T814_Spearman_PDO$estimate
      }
      
      if (T814_Pearson_PDO$p.value < 0.05) {
        T814_Pearson[r,c,2] = T814_Pearson_PDO$estimate
      }
      
      # NPGO
      T814_Spearman_NPGO <- cor.test(Thick814, NPGO_9524$NPGO,
                                     use = "pairwise.complete.obs", 
                                     method = "spearman")
      
      T814_Pearson_NPGO <- cor.test(Thick814, NPGO_9524$NPGO,
                                    use = "pairwise.complete.obs", 
                                    method = "pearson")
      
      if (T814_Spearman_NPGO$p.value < 0.05) {
        T814_Spearman[r,c,3] = T814_Spearman_NPGO$estimate
      }
      
      if (T814_Pearson_NPGO$p.value < 0.05) {
        T814_Pearson[r,c,3] = T814_Pearson_NPGO$estimate
      }
    }
    rm(Depth8, Thick814)
    rm(list = ls(pattern = "*_ONI"))
    rm(list = ls(pattern = "*_PDO"))
    rm(list = ls(pattern = "*_NPGO"))
  }
}
rm(r, c)
```

```{r, echo = FALSE}
# Make some variables to make making the table neater
# We don't have to do this, and it could create some errors, but I think it's worth it
C_Spear_min <- signif(apply(Catchability_Spearman, 3, min, na.rm = TRUE),2)
C_Spear_max <- signif(apply(Catchability_Spearman, 3, max, na.rm = TRUE),2)
C_Pear_min <- signif(apply(Catchability_Pearson, 3, min, na.rm = TRUE),2)
C_Pear_max <- signif(apply(Catchability_Pearson, 3, max, na.rm = TRUE),2)

O_Spear_min <- signif(apply(O2_Spearman, 3, min, na.rm = TRUE),2)
O_Spear_max <- signif(apply(O2_Spearman, 3, max, na.rm = TRUE),2)
O_Pear_min <- signif(apply(O2_Pearson, 3, min, na.rm = TRUE),2)
O_Pear_max <- signif(apply(O2_Pearson, 3, max, na.rm = TRUE),2)

D_Spear_min <- signif(apply(D8_Spearman, 3, min, na.rm = TRUE),2)
D_Spear_max <- signif(apply(D8_Spearman, 3, max, na.rm = TRUE),2)
D_Pear_min <- signif(apply(D8_Pearson, 3, min, na.rm = TRUE),2)
D_Pear_max <- signif(apply(D8_Pearson, 3, max, na.rm = TRUE),2)

T_Spear_min <- signif(apply(T814_Spearman, 3, min, na.rm = TRUE),2)
T_Spear_max <- signif(apply(T814_Spearman, 3, max, na.rm = TRUE),2)
T_Pear_min <- signif(apply(T814_Pearson, 3, min, na.rm = TRUE),2)
T_Pear_max <- signif(apply(T814_Pearson, 3, max, na.rm = TRUE),2)
```


Rough outline of my plan before I leave for the day:
get the min and max correlations for pixel-by-pixel corrs

Get the values for corrs with regional time serise

maybe load in some figures to help tell the story


### Results
Some text here.

|                           |               **ONI**                |              **PDO**                 |               **NPGO**               |
|:--------------------------|:------------------------------------:|:------------------------------------:|:------------------------------------:|
| **Catchability**          |                                      |                                      |                                      |
| *Spearman*                |`r C_Spear_min[1]`–`r C_Spear_max[1]` |`r C_Spear_min[2]`–`r C_Spear_max[2]` |`r C_Spear_min[3]`–`r C_Spear_max[3]` |
| *Pearson*                 |`r C_Pear_min[1]`–`r C_Pear_max[1]`   |`r C_Pear_min[2]`–`r C_Pear_max[2]`   |`r C_Pear_min[3]`–`r C_Pear_max[3]`   |
| **2 ml l^-^1 O~2~ Depth** |                                      |                                      |                                      |
| *Spearman*                |`r O_Spear_min[1]`–`r O_Spear_max[1]` |`r O_Spear_min[2]`–`r O_Spear_max[2]` |`r O_Spear_min[3]`–`r O_Spear_max[3]` |
| *Pearson*                 |`r O_Pear_min[1]`–`r O_Pear_max[1]`   |`r O_Pear_min[2]`–`r O_Pear_max[2]`   |`r O_Pear_min[3]`–`r O_Pear_max[3]`   |
| **8 °C Depth**            |                                      |                                      |                                      |
| *Spearman*                |`r D_Spear_min[1]`–`r D_Spear_max[1]` |`r D_Spear_min[2]`–`r D_Spear_max[2]` |`r D_Spear_min[3]`–`r D_Spear_max[3]` |
| *Pearson*                 |`r D_Pear_min[1]`–`r D_Pear_max[1]`   |`r D_Pear_min[2]`–`r D_Pear_max[2]`   |`r D_Pear_min[3]`–`r D_Pear_max[3]`   |
| **8–14 °C Thickness**     |                                      |                                      |                                      |
| *Spearman*                |`r T_Spear_min[1]`–`r T_Spear_max[1]` |`r T_Spear_min[2]`–`r T_Spear_max[2]` |`r T_Spear_min[3]`–`r T_Spear_max[3]` |
| *Pearson*                 |`r T_Pear_min[1]`–`r T_Pear_max[1]`   |`r T_Pear_min[2]`–`r T_Pear_max[2]`   |`r T_Pear_min[3]`–`r T_Pear_max[3]`   |

: Range of significant (*p* < 0.05) Spearman rank and Pearson correlations across the area of inters {#table_1}

Significant spatial correlations between oceanography are shown in @table_1.

Note to self: Map full area because it's helpful to understanding the gyre dynamics,
but in the discussion can include masked correlations to emphasize effect on fishery.
```{r, echo = FALSE}
# See what variables there are on each grid
# print(Catchability_nc$grid)
# 
# names(Catchability_nc$variable)
# 
# # Let's test out creating monthly averages, to see if I'm getting "group_by" correct
# Montly_catchability <- Catchability_t |>
#   group_by(as.numeric(Month)) |>
#   summarise(Monthly_Mean = mean(Catchability, na.rm = TRUE))
# 
# 
# Lat = ncread('../OceanData/BigeyeCatchability.nc', 'Latitude');
# Lon = ncread('../OceanData/BigeyeCatchability.nc', 'Longitude');
# Month = ncread('../OceanData/BigeyeCatchability.nc', 'Month');
# Catchability = ncread('../OceanData/BigeyeCatchability.nc', 'Catchability');
# GODAS_Depth = ncread('../OceanData/GODAS_1deg.nc', 'LEV1_31');
# GODAS = ncread('../OceanData/GODAS_1deg.nc', 'GODAS_REGRID');
# O2_2mlpl = ncread('../OceanData/O2_2mlpl_depth_1deg_noInterp.nc', 'GLORYS_O2_2MLPL_REGRID');
# ONI = readtable('../ClimateIndices/ONI_withPhases.csv');
# PDO = readtable('../ClimateIndices/PDO.csv');
# NPGO = readtable('../ClimateIndices/NPGO.csv');
# 
# 
# # Advanced spatial filtering with multiple conditions
# filtered_data <- nc_data %>%
#   hyper_filter(
#     lon = between(lon, 140, 220),           # Specific longitude range
#     lat = between(lat, -60, 0),             # Specific latitude range
#     time = between(time, "2000-01-01", "2020-12-31"),  # Time period
#     mask = !is.na(variable_name)              # Remove missing values
#   ) %>%
#   hyper_array(variable = "variable_name")
```


