---
title: "CPUEtable"
format: docx
editor: 
  markdown: 
    wrap: 72
---

# Purpose
The purpose of this script is to create a table that has the significant Pearson
correlations between each climate index and various fishery measures.  

I'm sure there are more efficient ways to do this, but I'm drawing on code that 
I wrote during a power outage (when I couldn't install packages or look things
up).  Since the code exists, I'm running with it.

This script borrows code from:
IndexAnalysis/SpeciesWithClimateIndicies.R (thanks, Johanna!)

```{r, echo = FALSE, message = FALSE}
# Set up environment
library(tidyverse, quietly = TRUE)
library(here)
library(tidync)
library(reshape2)
library(knitr)
```

```{r, echo = FALSE}
# Load and wrangle data
# when loading climate indices: 
# filter to period of interest, create common column names, and combine
oni <- read_csv(here("ClimateIndices", "ONI_withPhases.csv"), show_col_types = FALSE) |>
  filter(YR >= 1995 & YR <= 2024) |> 
  rename(Year=YR, Month=CentralMonth)
npgo <- read_csv(here("ClimateIndices", "NPGO.csv"), show_col_types = FALSE) |>
  filter(YEAR >= 1995 & YEAR <= 2024) |>
  rename(Year=YEAR, Month=MONTH)
pdo <- read_csv(here("ClimateIndices", "PDO.csv"), show_col_types = FALSE) |>
  filter(Year >= 1995 & Year <= 2024)
# join indices together
climIdx <- left_join(oni, pdo, by = join_by(Month, Year)) |>
  left_join(npgo, by = join_by(Month, Year)) |> 
  mutate(MonthIndex=1:360)

# Tidy up
rm(npgo, oni, pdo)

# Create tidync objects
Effort_nc <- tidync(here("FisheryData", "TotalEffort.nc"))
Bigeye_nc <- tidync(here("FisheryData", "TotalBigeyeCaught.nc"))
Yellowfin_nc <- tidync(here("FisheryData", "TotalYellowfinCaught.nc"))
Mahi_nc <- tidync(here("FisheryData", "TotalMahiCaught.nc"))
Pomfret_nc <- tidync(here("FisheryData", "TotalPomfretCaught.nc"))
Sword_nc <- tidync(here("FisheryData", "TotalSwordfishCaught.nc"))
PMUS_nc <- tidync(here("FisheryData", "TotalPMUS.nc"))

# Extract variables
# In a tidy-compatible format 
# and convert lat, lon, and month to numeric
Effort_t <- Effort_nc |>
  hyper_tibble() |>
  mutate_at(c('Longitude', 'Latitude', 'Month'), as.numeric)
Bigeye_t <- Bigeye_nc |>
  hyper_tibble() |>
  mutate_at(c('Longitude', 'Latitude', 'Month'), as.numeric)
Yellowfin_t <- Yellowfin_nc |>
  hyper_tibble() |>
  mutate_at(c('Longitude', 'Latitude', 'Month'), as.numeric)
Mahi_t <- Mahi_nc |>
  hyper_tibble() |>
  mutate_at(c('Longitude', 'Latitude', 'Month'), as.numeric)
Pomfret_t <- Pomfret_nc |>
  hyper_tibble() |>
  mutate_at(c('Longitude', 'Latitude', 'Month'), as.numeric)
Sword_t <- Sword_nc |>
  hyper_tibble() |>
  mutate_at(c('Longitude', 'Latitude', 'Month'), as.numeric)
PMUS_t <- PMUS_nc |>
  hyper_tibble() |>
  mutate_at(c('Longitude', 'Latitude', 'Month'), as.numeric)

# Sum effort and catch by month
Effort_TS <- Effort_t |>
  group_by(Month) |>
  summarise(Effort_ts = sum(`Total Effort`, na.rm = TRUE))
Bigeye_TS <- Bigeye_t |>
  group_by(Month) |>
  summarise(Bigeye_ts = sum(`Total Bigeye Caught`, na.rm = TRUE))
Yellowfin_TS <- Yellowfin_t |>
  group_by(Month) |>
  summarise(Yellowfin_ts = sum(`Total Yellowfin Caught`, na.rm = TRUE))
Mahi_TS <- Mahi_t |>
  group_by(Month) |>
  summarise(Mahi_ts = sum(`Total Mahi Caught`, na.rm = TRUE))
Pomfret_TS <- Pomfret_t |>
  group_by(Month) |>
  summarise(Pomfret_ts = sum(`Total Pomfret Caught`, na.rm = TRUE))
Sword_TS <- Sword_t |>
  group_by(Month) |>
  summarise(Sword_ts = sum(`Total Swordfish Caught`, na.rm = TRUE))
PMUS_TS <- PMUS_t |>
  group_by(Month) |>
  summarise(PMUS_ts = sum(`Total PMUS Caught`, na.rm = TRUE))
```

```{r, echo = FALSE}
# Now, build our big data frame that we're going to loop through for correlations
AllData <- data.frame(ONI = climIdx$ONI, 
                      PDO = climIdx$PDO, 
                      NPGO = climIdx$NPGO,
                      Effort = Effort_TS$Effort_ts,
                      BigeyeCPUE = Bigeye_TS$Bigeye_ts / Effort_TS$Effort_ts * 1000,
                      YellowfinCPUE = Yellowfin_TS$Yellowfin_ts / Effort_TS$Effort_ts * 1000,
                      MahiCPUE = Mahi_TS$Mahi_ts / Effort_TS$Effort_ts * 1000,
                      PomfretCPUE = Pomfret_TS$Pomfret_ts / Effort_TS$Effort_ts * 1000,
                      SwordCPUE = Sword_TS$Sword_ts / Effort_TS$Effort_ts * 1000,
                      PMUSCPUE = PMUS_TS$PMUS_ts / Effort_TS$Effort_ts * 1000)

# Tidy up
rm(list = setdiff(ls(), "AllData"))
```

```{r, echo = FALSE}
# Loop through and do the correlations
# Create empty matrix to fill
AllCors <- matrix(NA, nrow = 3, ncol = 7)
for (r in seq(1, 3, 1)) {
  for (c in seq(1, 7, 1)) {
    cor_hold <- cor.test(AllData[,r], AllData[,c+3])
    if (cor_hold$p.value < 0.05) {
      AllCors[r,c] <- cor_hold$estimate
    }
    c <- c+1
  }
  r <- r+1
}
rm(r, c)

rownames(AllCors) <- colnames(AllData)[1:3]
colnames(AllCors) <- colnames(AllData)[4:10]
```

```{r, echo = FALSE}
# Tidy some things up
AllCors <- as.data.frame(AllCors)

# Limit significant figures for plotting
AllCors <- signif(AllCors, 3)

# Create the table
kable(AllCors)
```
