---
title: "IndexTable"
format: docx
editor: 
  markdown: 
    wrap: 72
---

# Purpose

The purpose of this code is to generate a table that shows the range and
distribution of climate modes and effort across them.

This script borrows code from:
IndexAnalysis/SpeciesWithClimateIndicies.R (thanks, Johanna!)

```{r, echo = FALSE, message = FALSE}
# Set up environment
library(tidyverse, quietly = TRUE)
library(here)
library(tidync)
library(reshape2)
```

```{r, echo = FALSE}
# Load and wrangle data
# when loading climate indices: 
# filter to period of interest, create common column names, and combine
oni <- read_csv(here("ClimateIndices", "ONI_withPhases.csv"), show_col_types = FALSE) |>
  filter(YR >= 1995 & YR <= 2024) |> 
  rename(Year=YR, Month=CentralMonth)
npgo <- read_csv(here("ClimateIndices", "NPGO.csv"), show_col_types = FALSE) |>
  filter(YEAR >= 1995 & YEAR <= 2024) |>
  rename(Year=YEAR, Month=MONTH)
pdo <- read_csv(here("ClimateIndices", "PDO.csv"), show_col_types = FALSE) |>
  filter(Year >= 1995 & Year <= 2024)
# join indices together
climIdx <- left_join(oni, pdo, by = join_by(Month, Year)) |>
  left_join(npgo, by = join_by(Month, Year)) |> 
  mutate(MonthIndex=1:360)

# Limit the significant figures in the PDO and NPGO, for table crafting
climIdx$PDO <- signif(climIdx$PDO, 3)
climIdx$NPGO <- signif(climIdx$NPGO, 3)

# Tidy up
rm(npgo, oni, pdo)

# Create tidync objects
Effort_nc <- tidync(here("FisheryData", "TotalEffort.nc"))

# Extract variables
# Retaining netCDF dimensions
# Effort_a <- Effort_nc |>
#   hyper_array()
# In a tidy-compatible format
Effort_t <- Effort_nc |>
  hyper_tibble()

# Correct lat, lon, and month to be numbers instead of characters
Effort_t$Longitude <- as.numeric(Effort_t$Longitude)
Effort_t$Latitude <- as.numeric(Effort_t$Latitude)
Effort_t$Month <- as.numeric(Effort_t$Month)

# Sum effort by month
Effort_monthly <- Effort_t |>
  group_by(Month) |>
  summarise(EffortSummed = sum(`Total Effort`))
```

```{r, echo = FALSE}
# Create indices to create variables for the table
Nino <- which(!is.na(climIdx$ElNino))
Nina <- which(!is.na(climIdx$LaNina))
Neut <- which(!is.na(climIdx$Neutral))
Nada <- which(is.na(climIdx$ElNino) & is.na(climIdx$LaNina) & is.na(climIdx$Neutral))
PosPDO <- which(climIdx$PDO > 0)
NegPDO <- which(climIdx$PDO < 0)
PosNPGO <- which(climIdx$NPGO > 0)
NegNPGO <- which(climIdx$NPGO < 0)
TotEffort <- sum(Effort_monthly$EffortSummed)
```

Doing this as three tables to keep the line length reasonable. But they
can be combined after rendering. The dividing by 3.6 is a shortened
version of dividing by 360 and multiplying by 100 (total months,
transformed to %).  Something I learned: breaking the lines within the R call
preserves the table format, but breaking elsewhere affects formatting.

|        |  **ONI**  |  **ONI**  |  **ONI**  |  **ONI**  |
|:------:|:---------:|:---------:|:---------:|:---------:|
|        |   Pos     |    Neg    | Neutral   |    None   |
| Range  |`r min(climIdx$ONI[Nino])` - `r max(climIdx$ONI[Nino])` | `r 
min(climIdx$ONI[Nina])` - `r max(climIdx$ONI[Nina])` | `r 
min(climIdx$ONI[Neut])` - `r max(climIdx$ONI[Neut])` | `r 
min(climIdx$ONI[Nada])` - `r max(climIdx$ONI[Nada])` |
| Months |`r length(Nino)` (`r signif(length(Nino)/3.60,2)`%) | `r 
length(Nina)` (`r signif(length(Nina)/3.60,2)`%) |`r 
length(Neut)` (`r signif(length(Neut)/3.60,2)`%) | `r 
length(Nada)` (`r signif(length(Nada)/3.60,2)`%) |
| Effort |`r signif(sum(Effort_monthly$EffortSummed[Nino])/TotEffort*100,2)`% | `r signif(sum(Effort_monthly$EffortSummed[Nina])/TotEffort*100,2)`% | `r signif(sum(Effort_monthly$EffortSummed[Neut])/TotEffort*100,2)`% | `r signif(sum(Effort_monthly$EffortSummed[Nada])/TotEffort*100,2)`% |


|        |  **PDO**  | **PDO** |
|:------:|:---------:|:-------:|
|        |    Pos    |   Neg   |
| Range  |`r min(climIdx$PDO[PosPDO])` - `r max(climIdx$PDO[PosPDO])` | `r 
min(climIdx$PDO[NegPDO])` - `r max(climIdx$PDO[NegPDO])` |
| Months | `r length(PosPDO)` (`r signif(length(PosPDO)/3.60,2)`%) | `r 
length(NegPDO)` (`r signif(length(NegPDO)/3.60,2)`%) |
| Effort | `r signif(sum(Effort_monthly$EffortSummed[PosPDO])/TotEffort*100,2)`% | `r signif(sum(Effort_monthly$EffortSummed[NegPDO])/TotEffort*100,2)`% |
 
 
|        | **NPGO** | **NPGO** |
|:------:|:---------:|:-------:|
|        |    Pos    |   Neg   |
| Range  |`r min(climIdx$NPGO[PosNPGO])` - `r max(climIdx$NPGO[PosNPGO])` | `r 
min(climIdx$NPGO[NegNPGO])` - `r max(climIdx$NPGO[NegNPGO])` |
| Months | `r length(PosNPGO)` (`r signif(length(PosNPGO)/3.60,2)`%) | `r 
length(NegNPGO)` (`r signif(length(NegNPGO)/3.60,2)`%) |
| Effort | `r signif(sum(Effort_monthly$EffortSummed[PosNPGO])/TotEffort*100,2)`% | `r signif(sum(Effort_monthly$EffortSummed[NegNPGO])/TotEffort*100,2)`% |

