---
title: "LogbookAccess"
---

# Purpose
This script accesses logbook data and downloads the data of interest:  

* Begin Haul Date  
* Begin Haul Location
* Hooks per float (to determine deep-set vs. shallow-set)  
* Hooks per set
* Numbers caught and kept (for comparison, out of curiosity) of:
  * Bigeye
  * Mahi mahi
  * Yellowfin
  * Swordfish
  * Pomfret
* Bait type used
* Leader material (not available)

Thanks to Johanna Wren, upon whose climate indicator code this script is based!

```{r}
# Set up the environment
library(tidyverse, quietly = TRUE)
library(ROracle)
library(here)
```

# Access and download data 
Access data from Oracle using ROracle. 

```{r eval=FALSE}
# Put in your own credentials at the prompt. You have to have access to the dataset in question to be able to access it.
# Explore the data tables. This doesn't have to be done every time but it's helpful the first time and when trying to figure out what data are needed and available. 

# Establish connection with the database
con <- dbConnect(dbDriver("Oracle"),
                 username = rstudioapi::askForPassword("Oracle user name"),
                 password = rstudioapi::askForPassword("Oracle password"), 
                 dbname = "PIC")

# Name schema to use
shma_log <- 'LLDS'

# Look at what views are available in schema.
table_log <- dbListTables(con, schema=shma_log)
table_log

# Define table name
tname_hdr <- paste(tolower(shma_log),table_log[grep(paste0('LLDS_HDR_2025'), table_log)], sep='.')  # this is the HDR table, not the DETAIL
tname_detail <- paste(tolower(shma_log),table_log[grep(paste0('LLDS_DETAIL_2025'), table_log)], sep='.')

# Look at column names available in the table
tableInfo_hdr <- dbListFields(con, schema=shma_log, name=table_log[grep('LLDS_HDR_2025', table_log)])
tableInfo_detail <- dbListFields(con, schema=shma_log, name=table_log[grep('LLDS_DETAIL_2025', table_log)])
tableInfo_hdr
tableInfo_detail
```

```{r}
# Download the logbook data
# Query the database - hdr data
res <- dbSendQuery(con, paste("SELECT TRIPNUM, RSCH_EXPMTL_CODE, PERMITNUM, BAIT, HOOKSSET, MINHKSFLT, MAXHKSFLT, BHLATDEG, BHLATMIN, BHLONGDEG, BHLONGMIN, SERIALNUM, FLEET, BH_YR, BH_MON", 
                              "FROM", tname_hdr,
                              "WHERE BH_YR >= '1995' AND RSCH_EXPMTL_CODE IS NULL AND SERIALNUM IS NOT NULL AND FLEET = 'HI'")) 

logbookHDR <- dbFetch(res)

# Check data to make sure it's want we need
dim(logbookHDR)
head(logbookHDR)
str(logbookHDR)  # It looks like a lot of the numeric fields are read in as characters

# Convert character fields to numeric
logbookHDR <- logbookHDR %>% 
  mutate_at(c('TRIPNUM', 'PERMITNUM', 'BAIT', 'HOOKSSET', 'MINHKSFLT', 'MAXHKSFLT', 'BHLATDEG', 'BHLATMIN', 'BHLONGDEG', 'BHLONGMIN', 'SERIALNUM'), 
            as.numeric) 
str(logbookHDR)

# Save dataset (the internet appears to unanimously agree that rds is superior to csv in terms of speed, memory, metadata preservation, and more)
saveRDS(object = logbookHDR, file = here("FisheryData", "LogbookHDR.rds"))

# Clear query results
dbClearResult(res)

# Query the database - detail data
res <- dbSendQuery(con, paste("SELECT HDR_TRIPNUM, HDR_SERIALNUM, ENGLISH_NAME, SPECIES, NUMKEPT, NUMRELEASED, FLEET, BH_YR, BH_MON, PERMITNUM, HOOKSSET, MAXHKSFLT, RSCH_EXPMTL_CODE",
                              "FROM", tname_detail,
                              "WHERE BH_YR >= '1995' AND RSCH_EXPMTL_CODE IS NULL AND HDR_SERIALNUM IS NOT NULL AND FLEET = 'HI'")) 

logbookDETAIL <- dbFetch(res)

# Check data to make sure it's want we need
dim(logbookDETAIL)
head(logbookDETAIL)
str(logbookDETAIL)  # It looks like a lot of the numeric fields are read in as characters

# Convert character fields to numeric
logbookDETAIL <- logbookDETAIL %>% 
  mutate_at(c('HDR_TRIPNUM', 'HDR_SERIALNUM', 'SPECIES', 'PERMITNUM'), 
            as.numeric) 
str(logbookDETAIL)

# Save dataset (the internet appears to unanimously agree that rds is superior to csv in terms of speed, memory, metadata preservation, and more)
saveRDS(object = logbookDETAIL, file = here("FisheryData", "LogbookDETAIL.rds"))

# Clear your query results
dbClearResult(res)
```

```{r}
# Close connection
dbDisconnect(con)
```


