---
title: "LogbookCombine"
---

# Purpose
This script combines the logbook HDR and DETAIL data into a single set-level 
dataset.  The HDR data are per set.  The DETAIL data are per fish.  This script
sums the number of each species kept and released for each set (the comparison
of kept vs. released is just for my own personal curiosity).  

All this happens after sets are limited to deep-sets only: >= 15 hooks per float.

```{r}
# Set up the environment
library(tidyverse, quietly = TRUE)
library(here)
```

```{r}
# Load the data
HDR <- readRDS(here("FisheryData", "LogbookHDR.rds"))
DETAIL <- readRDS(here("FisheryData", "LogbookDETAIL.rds"))
```

```{r}
# Add a unique ID number to each set record, for troubleshooting
# HDR <- HDR |>
#   mutate(OBJECTID = seq(1, dim(HDR)[1], 1))
```


```{r}
# Remove records that are missing data that we need
# First remove the RSCH_EXPMTL_CODE columns, because we specifically selected ones that were empty
HDR <- HDR |>
  mutate(RSCH_EXPMTL_CODE = NULL)
DETAIL <- DETAIL |>
  mutate(RSCH_EXPMTL_CODE = NULL)

HDR_noNA <- na.omit(HDR)
DETAIL_noNA <- na.omit(DETAIL)

# Clean up
rm(HDR, DETAIL)
```

```{r}
# Filter to only deep sets
HDR_deep <- HDR_noNA |>
  filter(MINHKSFLT >= 15 | MAXHKSFLT >= 15)

DETAIL_deep <- DETAIL_noNA |>
  # filter(MAXHKSFLT >= 15) %>% # Taking out this condition because it's often 0 in early 1995
  mutate(HOOKSSET = as.numeric(HOOKSSET)) # for some reason HOOKSSET here is a character which causes issues later on

# Clean up
rm(HDR_noNA, DETAIL_noNA)
```

```{r}
# We're also going to combine the deg-min location info into decimal degrees and get rid of the old (annoying) columns
HDR_deep <- HDR_deep |>
  mutate(BH_LAT = BHLATDEG + BHLATMIN/60,
         BH_LON = BHLONGDEG + BHLONGMIN/60) %>% 
  select(-BHLATDEG:-BHLONGMIN)
```

```{r}
# Identify all the unique species caught
# Thankfully, this is logbook data so the list is short compared to observer data
# Also, order the species numbers
species_list <- sort(unique(DETAIL_deep$SPECIES))

# Okay, well, "short" was optimistic.  So.
# We'll just pull out our five species of interest.  
# On another day, when a clear need arises, this can be added to to include additional species
# BIGEYE TUNA: 16
# MAHIMAHI: 11
# YELLOWFIN TUNA: 17
# SWORDFISH: 6
# POMFRET: 21

# Adding in code to pull all PMUS species that are recorded in the logbook
# PMUS <- c(1, 2, 3, 4, 5, 6, 7, 8, 9,
#           11, 12, 13,
#           15, 16, 17, 18, 19, 20, 21, 22,
#           24, 25) #,
          # 37, 38) while these are listed as PMUS species, none of them are recorded
          # including them here and down below messes things up
          # they're 'DOGTOOTHTUNA' and 'KAWAKAWA', respectively
```

```{r}
# combine HDR and DETAIL
library(data.table)
# first, make detail "wide" instead of "long" for the NUMKEPT and NUMRELEASED columns for our five species
# This is easiest using data.table 
DETAIL_wide <- DETAIL_deep %>% 
  as.data.table() %>% 
  # filter(SPECIES %in% c(16,11,17,6,21)) %>% 
  # filter(SPECIES %in% PMUS) %>% 
  dcast(formula = HDR_SERIALNUM + HDR_TRIPNUM + PERMITNUM + BH_YR + BH_MON + HOOKSSET ~ SPECIES, value.var=c('NUMFINNED', 'NUMKEPT', 'NUMRELEASED'), fun.aggregate = sum)  %>% 
  # dcast(formula = HDR_SERIALNUM + HDR_TRIPNUM + PERMITNUM + BH_YR + BH_MON + HOOKSSET + MAXHKSFLT ~ SPECIES, value.var=c('NUMKEPT', 'NUMRELEASED'), fun.aggregate = sum) %>% 
  rename(SERIALNUM=HDR_SERIALNUM, TRIPNUM=HDR_TRIPNUM)

# Combine the kept&released data with the HDR data
DeepSets <- merge.data.table(as.data.table(HDR_deep), DETAIL_wide)
# Rename columns (they are ordered by increasing number!). Take care here to make sure you are renaming them correctly
# When adding the OBJECTID field, the columns move over one (was 13:22, 13:56)
# setnames(DeepSets, c(14:23), paste(rep(c('SWORDFISH', 'MAHI', 'BIGEYE', 'YELLOWFIN', 'POMFRET'), 2), rep(c('KEPT', 'RELEASED'), each=5), sep='_'))
# setnames(DeepSets, c(14:57), paste(rep(c('BLUEMARLIN', 'STRIPEDMARLIN', 'BLACKMARLIN',
#                                           'SAILFISH', 'SPEARFISH', 'SWORDFISH',
#                                           'BLUESHARK', 'MAKOSHARK', 'THRESHERSHARK',
#                                           'MAHI', 'OPAH', 'WAHOO',
#                                           'ALBACORE', 'BIGEYE', 'YELLOWFIN', 'OTHERTUNA', 'BLUEFIN',
#                                           'OILIFISH', 'POMFRET',
#                                           'SKIPJACK',
#                                           'WHITETIPSHARK', 'SILKYSHARK'), 2),
#                                    rep(c('KEPT', 'RELEASED'), each = 22), sep = '_'))
DeepSets <- as.data.frame(DeepSets)
```

```{r}
# Double-check that we're not including any shallow sets
maybe_shallow_minhks <- which(DeepSets$MINHKSFLT > 0 & DeepSets$MINHKSFLT < 15) # 31 entries, 32 for PMUS and all spp versions
maybe_shallow_maxhks <- which(DeepSets$MAXHKSFLT > 0 & DeepSets$MAXHKSFLT < 15) # empty
definitely_shallow <- which(DeepSets$MINHKSFLT < 15 & DeepSets$MAXHKSFLT == 0) # empty

# Remove potential shallow sets, just to be cautious
DeepSets <- DeepSets[-maybe_shallow_minhks,]
```

```{r}
# Save output in two formats for different platforms
saveRDS(object = DeepSets, file = here("FisheryData", "DeepSets_AllSpecies.rds"))
write_csv(DeepSets, file = here("FisheryData", "DeepSets_AllSpecies.csv"))
```





