---
title: "LogbookCombine"
---

# Purpose
This script combines the logbook HDR and DETAIL data into a single set-level 
dataset.  The HDR data are per set.  The DETAIL data are per fish.  This script
sums the number of each species kept and released for each set (the comparison
of kept vs. released is just for my own personal curiosity).  

All this happens after sets are limited to deep-sets only: >= 15 hooks per float.

```{r}
# Set up the environment
library(tidyverse, quietly = TRUE)
library(here)
```

```{r}
# Load the data
HDR <- readRDS(here("FisheryData", "LogbookHDR.rds"))
DETAIL <- readRDS(here("FisheryData", "LogbookDETAIL.rds"))
```

```{r}
# Remove records that are missing data that we need
# First remove the RSCH_EXPMTL_CODE columns, because we specifically selected ones that were empty
HDR <- HDR |>
  mutate(RSCH_EXPMTL_CODE = NULL)
DETAIL <- DETAIL |>
  mutate(RSCH_EXPMTL_CODE = NULL)

HDR_noNA <- na.omit(HDR)
DETAIL_noNA <- na.omit(DETAIL)

# Clean up
rm(HDR, DETAIL)
```

```{r}
# Filter to only deep sets
HDR_deep <- HDR_noNA |>
  filter(MINHKSFLT >= 15 & MAXHKSFLT >= 15)

DETAIL_deep <- DETAIL_noNA |>
  filter(MAXHKSFLT >= 15)

# Clean up
rm(HDR_noNA, DETAIL_noNA)
```

```{r}
# We're also going to combine the deg-min location info into decimal degrees
HDR_deep <- HDR_deep |>
  mutate(BH_LAT = BHLATDEG + BHLATMIN/60,
         BH_LON = BHLONGDEG + BHLONGMIN/60)

# And get rid of the old (annoying) columns
# We could probably do this together with the previous step
HDR_deep <- HDR_deep |>
  mutate(BHLATDEG = NULL,
         BHLATMIN = NULL,
         BHLONGDEG = NULL,
         BHLONGMIN = NULL)
```

```{r}
# Identify all the unique species caught
# Thankfully, this is logbook data so the list is short compared to observer data
# Also, order the species numbers
species_list <- sort(unique(DETAIL_deep$SPECIES))

# Okay, well, "short" was optimistic.  So.
# We'll just pull out our five species of interest.  
# On another day, when a clear need arises, this can be added to to include additional species
# BIGEYE TUNA: 16
# MAHIMAHI: 11
# YELLOWFIN TUNA: 17
# SWORDFISH: 6
# POMFRET: 21
```

```{r}
# Create our new dataset, including adding new column
DeepSets <- HDR_deep |>
  mutate(BIGEYE_KEPT = as.numeric(NA), 
         BIGEYE_RELEASED = as.numeric(NA),
         MAHI_KEPT = as.numeric(NA),
         MAHI_RELEASED = as.numeric(NA),
         YELLOWFIN_KEPT = as.numeric(NA),
         YELLOWFIN_RELEASED = as.numeric(NA),
         SWORDFISH_KEPT = as.numeric(NA),
         SWORDFISH_RELEASED = as.numeric(NA),
         POMFRET_KEPT = as.numeric(NA), 
         POMFRET_RELEASED = as.numeric(NA))

# Loop through all the sets
# This is agonizingly slow, but it's the least-slow approach I could make work
for (s in seq(1, dim(HDR_deep)[1], 1)) {
  # Match the identifying details, some of which restart each year, sigh...
  fish_idx <- which(DETAIL_deep$HDR_SERIALNUM == DeepSets$SERIALNUM[s] & 
                    DETAIL_deep$HDR_TRIPNUM == DeepSets$TRIPNUM[s] &
                    DETAIL_deep$PERMITNUM == DeepSets$PERMITNUM[s] & 
                    DETAIL_deep$BH_YR == DeepSets$BH_YR[s] & 
                    DETAIL_deep$BH_MON == DeepSets$BH_MON[s] &
                    DETAIL_deep$HOOKSSET == DeepSets$HOOKSSET[s] &
                    DETAIL_deep$MAXHKSFLT == DeepSets$MAXHKSFLT[s])
  fish_hold <- DETAIL_deep[fish_idx,]
  
  # Sum number kept and released for each species
  bigeye_idx <- which(fish_hold$SPECIES == 16)
  mahi_idx <- which(fish_hold$SPECIES == 11)
  yellowfin_idx <- which(fish_hold$SPECIES == 17)
  swordfish_idx <- which(fish_hold$SPECIES == 6)
  pomfret_idx <- which(fish_hold$SPECIES == 21)
  
  DeepSets$BIGEYE_KEPT[s] <- sum(fish_hold$NUMKEPT[bigeye_idx])
  DeepSets$MAHI_KEPT[s] <- sum(fish_hold$NUMKEPT[mahi_idx])
  DeepSets$YELLOWFIN_KEPT[s] <- sum(fish_hold$NUMKEPT[yellowfin_idx])
  DeepSets$SWORDFISH_KEPT[s] <- sum(fish_hold$NUMKEPT[swordfish_idx])
  DeepSets$POMFRET_KEPT[s] <- sum(fish_hold$NUMKEPT[pomfret_idx])
  
  DeepSets$BIGEYE_RELEASED[s] <- sum(fish_hold$NUMRELEASED[bigeye_idx])
  DeepSets$MAHI_RELEASED[s] <- sum(fish_hold$NUMRELEASED[mahi_idx])
  DeepSets$YELLOWFIN_RELEASED[s] <- sum(fish_hold$NUMRELEASED[yellowfin_idx])
  DeepSets$SWORDFISH_RELEASED[s] <- sum(fish_hold$NUMRELEASED[swordfish_idx])
  DeepSets$POMFRET_RELEASED[s] <- sum(fish_hold$NUMRELEASED[pomfret_idx])
  
  # Tidy up
  rm(fish_hold)
  rm(list=ls(pattern="_idx"))
}
rm(s)
```

```{r}
# Save output in two formats for different platforms
saveRDS(object = DeepSets, file = here("FisheryData", "DeepSets.rds"))
write_csv(DeepSets, file = here("FisheryData", "DeepSets.csv"))
```





